---
title: "NWUNEMELU .C.S"
author: "S.C. Nwunemelu"
date: "3/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)### Read Excel Files. Import excel files into R

library(tidyverse)###Tidyverse Is the collection of essential R packages for data science

```

##In this section i call in my two datasets, view my datasets and also the names of the ##two variables in the datasets with the number of rows each have, i find out that two 
## datasets has same variables but different numbers of observations . so as to know how to merge the two dataset together.

```{r}
###Step1: Data Preprocessing and Summarizing Datasets

hotel1 <- read.csv("C:\\Users\\User\\Desktop\\chibu\\H1.csv")

hotel2 <- read.csv("C:\\Users\\User\\Desktop\\chibu\\H2.csv")

str(hotel1)

View(hotel1)

View(hotel2)

names(hotel1)

names(hotel2)

str(hotel2)



```

###Merging my two datasets together using rowbind. I choose row bind because the two datasets has the same variables but different numbers of observation.
```{r}
Hoteldata <- rbind(hotel1,hotel2)

View(Hoteldata)

str(Hoteldata)

summary(Hoteldata)

###In this part i was able to combine my two datasets together, view in

## the data sets i observe that they were alot of null values in the 


##AGENT and COMPANY columns,using str() function to check the 

##datatype of the datasets and also summary() function to check for 

##NA'S in the dataset




```

```{r}
### from my observation i found out that AGENT and COMPANY has too many null values in the columns

Hoteldata$Agent = NULL###This function removes the AGENT variable from the datasets

Hoteldata$Company = NULL

Hoteldata$ReservationStatusDate = NULL

Hoteldata$Country = NULL

Hoteldata$ReservationStatus= NULL

names(Hoteldata)

###I use the name()function to confirm the change i made on my ##datasets.

```

```{r}
## To replace NA'S in the dataset with median

sum(is.na(Hoteldata))#To check the number of NA'S in the dataset 
## which is 4

names(which(colSums(is.na(Hoteldata))>0)) ##This Function shows the variable with NA'S

##Replace Missing Values with median

Hoteldata$Children[is.na(Hoteldata$Children)] <- median(Hoteldata$Children,na.rm =TRUE)

sum(is.na(Hoteldata)) ##To check if the NA'S has been removed

summary(Hoteldata)

```

```{r}
###Converting variables datatype

Hoteldata$IsCanceled <- as.factor(Hoteldata$IsCanceled)

Hoteldata$Adults <- as.integer(Hoteldata$Adults)

Hoteldata$ArrivalDateDayOfMonth <- as.integer(Hoteldata$ArrivalDateDayOfMonth)

Hoteldata$ArrivalDateMonth <- as.factor(Hoteldata$ArrivalDateMonth)

Hoteldata$ArrivalDateWeekNumber <- as.integer(Hoteldata$ArrivalDateWeekNumber)

Hoteldata$ArrivalDateYear <- as.integer(Hoteldata$ArrivalDateYear)

```

```{r}
Hoteldata$AssignedRoomType <- as.factor(Hoteldata$AssignedRoomType)

Hoteldata$Babies <- as.integer(Hoteldata$Babies)

Hoteldata$BookingChanges<- as.integer(Hoteldata$BookingChanges)

Hoteldata$Children <- as.integer(Hoteldata$Children)

Hoteldata$CustomerType <- as.factor(Hoteldata$CustomerType)

Hoteldata$DaysInWaitingList <- as.integer(Hoteldata$DaysInWaitingList)

```

```{r}
Hoteldata$DepositType <- as.factor(Hoteldata$DepositType)

Hoteldata$ReservedRoomType <- as.factor(Hoteldata$ReservedRoomType)

Hoteldata$StaysInWeekendNights <- as.integer(Hoteldata$StaysInWeekendNights)

Hoteldata$StaysInWeekNights <- as.integer(Hoteldata$StaysInWeekNights)

Hoteldata$TotalOfSpecialRequests <- as.integer(Hoteldata$TotalOfSpecialRequests)

Hoteldata$DistributionChannel<- as.factor(Hoteldata$DistributionChannel)


Hoteldata$IsRepeatedGuest<- as.factor(Hoteldata$IsRepeatedGuest)

```

```{r}
Hoteldata$LeadTime<- as.integer(Hoteldata$LeadTime)

Hoteldata$MarketSegment <- as.factor(Hoteldata$MarketSegment)

Hoteldata$Meal <- as.factor(Hoteldata$Meal)

Hoteldata$PreviousBookingsNotCanceled <- as.integer(Hoteldata$PreviousBookingsNotCanceled)

Hoteldata$PreviousCancellations <- as.integer(Hoteldata$PreviousCancellations)

Hoteldata$RequiredCarParkingSpaces <- as.integer(Hoteldata$RequiredCarParkingSpaces)

Hoteldata$ADR <- as.numeric(Hoteldata$ADR)

str(Hoteldata)
```

```{r}
install.packages("ggplot2")

install.packages("ggpbur")
install.packages("plotrix")
install.packages("stringr")

```

```{r}

library(tidyverse)

library(ggplot2)###ggplot2 is an R package dedicated to data visualization. It can greatly ##improve the quality and aesthetics of your graphics, 

library(plotrix)##The plotrix package is used to provide a method for getting many sorts of specialized plots ... 

library(stringr)##s a set of simple wrappers that make R's string functions more consistent, simpler and easier to use. 

library(ggpubr)##The ggpubr R package facilitates the creation of beautiful ggplot2-based graphs for researcher with non-advanced programming backgrounds.
```

```{r}
##I check to see the percentage of Booking is canceled and Not canceled

glimpse(Hoteldata$IsCanceled)

prop.table(table(Hoteldata$IsCanceled))##to check the percentage of the BOOKING is CANCELED ## OVER NOT CANCELED to know which one is greater than each other.
table(Hoteldata$IsCanceled)

```





```{r}
ggplot(Hoteldata,aes(x =ArrivalDateMonth, fill =IsCanceled))+
  theme_bw() +
  geom_bar() +
  coord_flip()+
  labs(y = "ADR")
title = "Hoteldata IsCanceled rate at which ArrivalDatemonth Vs ADR"

```

```{r}
Hoteldata %>%
  ggplot(aes(MarketSegment,fill = IsCanceled))+
  theme_bw()+
  geom_bar()+
  coord_flip()+
  labs(y="ArrivalDateYear")
 title = "Hoteldata IsCanceled rate at which MarketSegment Vs ArrivalDateYear" 
  
  
```

```{r}
Hoteldata %>%
  ggplot(aes(DepositType,fill = IsCanceled))+
  theme_bw()+
  geom_bar()+
  labs(y="DaysInWaitingList")
 title = "Hoteldata IsCanceled rate at which DaysInWaitingList Vs DepositType" 
  

```

```{r}
Hoteldata %>%
  ggplot(aes(AssignedRoomType,fill = IsCanceled))+
  theme_bw()+
  geom_bar()+
  
  labs(y="BookingChanges")
 title = "Hoteldata IsCanceled rate at which Assignedroom Vs BookingChanges" 
  
```

```{r}
Hoteldata %>%
  ggplot(aes(StaysInWeekendNights,fill = IsCanceled))+
  theme_bw()+
  geom_bar()+
  labs(y="DistributionChannel")
 title = "Hoteldata IsCanceled rate at which StaysInWeekendNights Vs DistributionChannel" 
  

```

```{r}
Hoteldata %>%
  ggplot(aes(CustomerType,fill = IsCanceled))+
  theme_bw()+
  geom_bar()+
  labs(y="LeadTime")
 title = "Hoteldata IsCanceled rate at which CustomerType Vs LeadTime" 
  
```

```{r}
Hoteldata %>%
  ggplot(aes(IsRepeatedGuest,fill = IsCanceled))+
  theme_bw()+
  geom_bar()+
  
  labs(y="TotalofSpecialRequests")
 title = "Hoteldata IsCanceled rate at which IsRepeatedGuest Vs TotalofSpecialRequests" 
  
```

```{r}
install.packages("PerformanceAnalytics")
install.packages("caret")
install.packages("e1071")
install.packages("randomForest")
install.packages("catools")

```

```{r}
library(PerformanceAnalytics)###provides an R package of econometric functions for ##performance and risk analysis of financial instruments.

library(caret)###(short for classification and Regression Training) contains functions to ##streamline the model training process for complex regression and classification problems.

library(e1071)###it provides a training function with standard and formula interfaces, and a predict() method.

library(randomForest)###The R package "randomForest" is used to create random forests.

library(caTools) ##
```

```{r}

library(randomForest)

library(caret)

str(Hoteldata)
```
```{r}
gc()
memory.limit(20000)

```

```{r}
###LOGISTIC REGRESSION MODEL

##Feature Selection

### using  random forest to get  the important  variables

memory.limit(10000)

output.forestHoteldata <- randomForest(IsCanceled ~., data = Hoteldata)

output.forestHoteldata

randomForest:: importance(output.forestHoteldata)


```

```{r}
order()
randomF_DF <- data.frame(Variables=row.names(output.forestHoteldata),
                         MeanDecreaseGini = output.forestHoteldata[,1])
randomF_DF <- randomF_DF[order(randomF_DF$MeanDecreaseGini,decrease=TRUE),]
randomF_DF
```

```{r}

###Creating training and testing data sets
##Spliting my datasets

view(Hoteldata)

Hotel_1 <- Hoteldata [which(Hoteldata$IsCanceled == 1),]

Hotel_2 <- Hoteldata[which(Hoteldata$IsCanceled == 0),]

view(Hotel_1)

view(Hotel_2)


```

```{r}

 set.seed(120)### as the machine to pick a random number
#### Set training data to 80% of the dataset (Balance my 1 and 0 values)

trainHoteldata_1 <- sample(1:nrow(Hotel_1),0.80*nrow(Hotel_1))### this sample out 80% of the 1's outcome 

trainHoteldata_2 <- sample(1:nrow(Hotel_2),0.80*nrow(Hotel_2))

view(trainHoteldata_1)### it will give you the indexes  of the 80% of 1's nrow

view(trainHoteldata_2)

```

```{r}
##put both in dataframe

trainingH_1 <- Hotel_1[trainHoteldata_1,] ### subsetting 

trainingH_1

trainingH_2 <- Hotel_2[trainHoteldata_2,]

trainingH_2
```

```{r}
trainHdata <- rbind(trainingH_1,trainingH_2)## combine the 1's & 0's together

view(trainHdata)
```

```{r}
####Creating testing datasets 

testHoteldata1 <-Hotel_1 [-trainHoteldata_1,]

testHoteldata2 <- Hotel_2[-trainHoteldata_2,]

view(testHoteldata1)

view(testHoteldata2)
```

```{r}
##Binding the two datasets together

testH <- rbind(testHoteldata1, testHoteldata2)
testH

view(testH)
```

```{r}
view(Hoteldata)
str(Hoteldata)
head(Hoteldata)

```

```{r}
###Creating a Model
logisticsmodelH <- glm(IsCanceled~.,data = trainHdata, family = binomial(link = "logit"))

summary(logisticsmodelH)
```

```{r}
logisticsmodelH_1 <- glm(IsCanceled~LeadTime+ArrivalDateYear+ArrivalDateMonth
                                            +StaysInWeekendNights+StaysInWeekNights
                                            +Adults+Children+Babies+Meal
                                            +MarketSegment+DistributionChannel
                                            +IsRepeatedGuest+PreviousCancellations
                                            +PreviousBookingsNotCanceled
                                            +ReservedRoomType+AssignedRoomType
                                            +BookingChanges+DepositType
                                            +DaysInWaitingList+CustomerType
                                            +ADR+RequiredCarParkingSpaces
                                            +TotalOfSpecialRequests,data =trainHdata,
                                             binomial(link = "logit"))
summary(logisticsmodelH_1)
```

```{r}
logisticsmodelH_2 <- glm(IsCanceled~.-RequiredCarParkingSpaces, data = trainHdata,
                         binomial(link="logit"))

summary(logisticsmodelH_2)

```

```{r}
logisticsmodelH_3 <- glm(IsCanceled~.-DaysInWaitingList, data = trainHdata,
                         binomial(link="logit"))

summary(logisticsmodelH_3)

```

```{r}
logisticsmodelH_4 <- glm(IsCanceled~LeadTime+ArrivalDateYear+ArrivalDateMonth
                                            +StaysInWeekendNights+StaysInWeekNights
                                            +Adults+Children+Babies
                                            +MarketSegment+DistributionChannel
                                            +IsRepeatedGuest+PreviousCancellations
                                            +PreviousBookingsNotCanceled
                                            +ReservedRoomType+AssignedRoomType
                                            +DepositType+CustomerType+ADR
                                            +RequiredCarParkingSpaces
                                            +TotalOfSpecialRequests,data =trainHdata,
                                             binomial(link = "logit"))
summary(logisticsmodelH_4)
```

```{r}
logisticsmodelH_5 <- glm(IsCanceled~LeadTime+ArrivalDateYear+ArrivalDateMonth
                                            
                                            +Adults+Children+Babies+Meal
                                            +MarketSegment
                                            +IsRepeatedGuest+PreviousCancellations
                                            +PreviousBookingsNotCanceled
                                            +ReservedRoomType+AssignedRoomType
                                            +BookingChanges+DepositType
                                          +CustomerType+ADR+RequiredCarParkingSpaces,
                                           data =trainHdata,
                                             binomial(link = "logit"))
summary(logisticsmodelH_5)
```

```{r}
##predict scores and add it to the testing data

predicted <- predict(logisticsmodelH,testH)
predicted

view(predicted)

```



```{r}
###Joining of the predicted variable to the dataset table

testH$predicted <- predicted

view(testH)

```

```{r}
##Analysis of Logistics Model and Model fit
modelsummaryH <- summary(logisticsmodelH)


modelsummaryH$coefficient



```

```{r}


AIC(logisticsmodelH)##penalty  for adding additional variables to the model

BIC(logisticsmodelH)### a stronger penalty for adding additional variables to the model
```

```{r}
###Test and Graph the model
##plot of Booking is canceled vs not canceled IsCanceled by Leadtime
library(ggplot2)
ggplot(testH,aes(x=IsCanceled,y=LeadTime))+geom_point()+stat_smooth(method="glm",se=FALSE)

###Same In Bar Graph
par(mfrow=c(1,1))
plot(IsCanceled~LeadTime,data=trainHdata)

```

```{r}
### Calling in my dataset
View(Hoteldata)
str(Hoteldata)
```

```{r}
##Scoring the values back on the original dataset
pred_H <- predict(logisticsmodelH,testH,"link")
head(pred_H)

```

```{r}


###Create new class prediction column on the original dataset for side by side comparison

testH$class <-ifelse(pred_H>0.5, 1,0)

view(testH)


```

```{r}

table(pred_H)

table(testH$IsCanceled)

```

```{r}
table(ActualValue =testH$IsCanceled,PredictedValue = pred_H>0.5)

```



```{r}
###foR ACCURACY
##(TN+FP)/(TN+FN+FP+TP)
(14722+4249)/(14722+4596+312+4249)
```

```{r}
###Create new class prediction column on the original dataset for side by side comparison
testH$class <-ifelse(pred>0.5, 1,0)
view(testH)

```



```{r}
###Practical REAL Time Prediction

###To predict if a customer Booking is canceled or Not


M <- data.frame(LeadTime=7,ArrivalDateYear=2015,ArrivalDateMonth= "July",
               ArrivalDateWeekNumber=27,ArrivalDateDayOfMonth=1,StaysInWeekendNights=0,
               StaysInWeekNights=4,Adults=2,Children=1,Babies=0,Meal="HB",           
               MarketSegment="Direct", DistributionChannel="Direct",IsRepeatedGuest=0,
               PreviousCancellations=0,PreviousBookingsNotCanceled=0,ReservedRoomType="C",
               AssignedRoomType="C",BookingChanges=0,DepositType="No Deposit",
               DaysInWaitingList=0,CustomerType="Transient",ADR=103.00,
               RequiredCarParkingSpaces=0,TotalOfSpecialRequests=1)

p <- predict(logisticsmodelH_1,M)

p

M$class <- ifelse(p>0,1,0)
M
```

```{r}
###Starting My RANDOM FOREST MODEL
Hoteldata$IsCanceled <- ifelse(Hoteldata$IsCanceled==0, 1,0)

View(Hoteldata)

Hoteldata$IsCanceled <- as.factor(Hoteldata$IsCanceled)

str(Hoteldata)
```

```{r}
##get predictors values by fitting another model this helps  to know important predictors

output.forestHoteldata <-randomForest(IsCanceled~.,data = Hoteldata)

##View the result
print(output.forestHoteldata)


### To check the Importance variable
randomForest:: importance(output.forestHoteldata)
 
```

```{r}
###Plotting a variable Importance graph
varImpPlot(output.forestHoteldata, las=2, col= "blue")
```

```{r}
###To check  the model Error at each number of the trees
plot(output.forestHoteldata)

```

```{r}
### Divide my dataset into training and testing datasets

set.seed(150)

split_tagHoteldata <-sample(nrow(Hoteldata),0.80*nrow(Hoteldata),replace = FALSE)

trainHoteldata <- Hoteldata[split_tagHoteldata,]

testHoteldata <- Hoteldata[-split_tagHoteldata,]

nrow(trainHoteldata)

nrow(testHoteldata)
 
summary(trainHoteldata)

summary(testHoteldata)
```

```{r}
view(trainHoteldata)
view(testHoteldata)
str(Hoteldata)
Hoteldata$IsCanceled <- as.factor(Hoteldata$IsCanceled)

```

```{r}

### Implementing a Random Forest Model to the dataset and also checking the importance

modelHoteldata_1 <- randomForest(IsCanceled~.,data = trainHoteldata, importance = TRUE)

  
modelHoteldata_1

##higher values in MeanDecreaseGini is the greater importance


importance(modelHoteldata_1)




```

```{r}
###Fine tune the random Forest Random
modelHoteldata_2 <- randomForest(IsCanceled~.,data = trainHoteldata,
                                            ntree=700,mtry=3,importance = TRUE)
    

modelHoteldata_2


```

```{r}
##predict on the train data

predict_trainHoteldata <- predict(modelHoteldata_1,trainHoteldata,type = "class")
predict_trainHoteldata
table(predict_trainHoteldata,trainHoteldata$IsCanceled)

```

```{r}
confusionMatrix(table(predict_testHoteldata,testHoteldata$IsCanceled))


```

```{r}
###Check the variable of importance


randomForest::importance(modelHoteldata_2)

varImpPlot(modelHoteldata_2,las=2,col='blue')

```

```{r}
##Predicting on Testing dataset
predtest <- predict(modelHoteldata_1, testHoteldata, type ="class")
predtest



```

```{r}
###Checking classification accuracy

mean(predtest == testHoteldata$IsCanceled)

table(predtest,testHoteldata$IsCanceled)
```

```{r}
confusionMatrix(table(predtest,testHoteldata$IsCanceled))
```

#DECISION TREE MODEL

```{r}
###install other Enhanced Tree plotters
install.packages("rpart")
install.packages("rpart.plot")
install.packages("rattle")
install.packages("RColorBrewer")
install.packages("tree")

install.packages("e1071")

install.packages("caTools")

```

```{r}
library(rpart)##it is using in building a classification or regression models of a very ##general structure using a two stage procedure.

library(rpart.plot)##it automatically scales and adjusts the displayed tree for best fit.

library(rattle)##It provides statistical and visual summaries of data, transform data so it ##can be readily modelled.

library(RColorBrewer)## is an R package that contains a ready to use color palettes for 
## creating beautiful graphics.

library(tree)##
library(caTools)
library(e1071)
library(caret)
```


##Building the decision Tree Model

```{r}

treemodel_H <- tree(IsCanceled~.,data=Hoteldata)
summary(treemodel_H)
##the Residual mean deviance 0.89% 

###the number terminal node is 6
###the misclassification error is 0.19% which is 19% is low which means the model is ok
### 5 variables were used for the construction of the trees.
```

```{r}
###plot the decision tree
par(mfrow=c(1,1))
plot(treemodel_H)
text(treemodel_H,pretty= 0)

```

```{r}
###plot the tree with repart
treemodel_H2 <- rpart(IsCanceled~., data = Hoteldata)##Using repart to create a model
summary(treemodel_H2)

```

```{r}
fancyRpartPlot(treemodel_H2)
```

```{r}
printcp(treemodel_H2)


```

```{r}
ptreeH <- prune(treemodel_H2,+ treemodel_H2$cptable[which.min(treemodel_H2$cptable
                                                     [,"xerror"]),"CP"])
summary(ptreeH)



```

```{r}
fancyRpartPlot(ptreeH,uniform = TRUE, main = 'Pruned Classification Tree')

```

```{r}
sample.split(Hoteldata$IsCanceled,SplitRatio = 0.70)->split_tag ###you have to sample 70% and assign to split_tag
split_tag
```

```{r}
subset(Hoteldata,split_tag==T)-> trainHoteldata ###contain the 70% train
trainHoteldata
nrow(trainHoteldata)
view(trainHoteldata)

```

```{r}
subset(Hoteldata,split_tag==F)-> testHoteldata ###contain the 30% test
testHoteldata
nrow(testHoteldata)

```

```{r}
###Building the model on the train data
### first fit an unpruned classification tree using all of the predictors
treemodel1_H = tree(IsCanceled~.,data = trainHoteldata)
summary(treemodel1_H)

```

```{r}
plot(treemodel1_H)
text(treemodel1_H,pretty = 0)
title(main = "Unpruned Classification Tree")
```

```{r}
##make Prediction with the test data.
tree_predH = predict(treemodel1_H,testHoteldata,type="class")
tree_predH
view(tree_predH)

```

```{r}
confusionMatrix(tree_predH,testHoteldata$IsCanceled)
```

```{r}
##Check Actual against the predicted Response
predresult <- cbind(testHoteldata,tree_predH)
view(predresult)
```

```{r}
#perform Cross validation
cv_tree=cv.tree(treemodel1_H,FUN = prune.misclass)
cv_tree ## 7 terminal nodes have the least deviation of 16603
###so in the next line we will prune
#this no 7 can vary

```

```{r}
pruned_tree =prune.misclass(treemodel_H,best=7)
plot(pruned_tree)
text(pruned_tree,pretty=0)
title(main="Cross Validated Prune Classification Tree")
```

```{r}
###Predict with the crossvalidated Pruned_tree
tree_predictH =predict(pruned_tree,testHoteldata,type = "class")
table(testHoteldata$IsCanceled,tree_predH)

```

```{r}
confusionMatrix(tree_predH,testHoteldata$IsCanceled)
confusionMatrix(tree_predictH,testHoteldata$IsCanceled)
```

```{r}
 
###get the accuracy
(8140+20863)/(8140+20863+1687+5127)
```
